---
title: "Homework 2"
author: "Alberto Pineda"
date: "2023-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
Create a vector of monthly returns from holding two stocks. Call the first vector stock1 containing only the values of the returns and
not the month names. Returns for the first nine months are:
a)
January: +$2000 February: +$4000 March: -$500 April: -$1000 May: +$7000 June: +$7500 July: -$800 August: +$5500 September: -$1000
b) Call the second vector stock2 containing only the values of the returns and
not the month names. Returns for the first nine months are:
January: +$3000 February: -$400 March: -$500 April: +$1000 May: -$700 June: +$7500 July: +$1250 August: +$6500 September: +$1000
c) Create a vector called months for storing month names - January, Febru-
ary,..., September.
d) Assign names of the months to the stock1 and stock2 vectors.
e) Create a vector total monthly to hold the total monthly gains or losses
from holding both stocks.
f)what is the total gain/loss over the last nine months from both the stocks?
Store this in a vector named total nine months
g) Check to see if the total returns from stock1 exceeds total returns from
stock2
h) Store the return of stock1 for March in stock1 March
i) Store the return of stock2 for September in stock2 Sept.
j) Store the returns from the second quarter (April, May, June) for stock1.
Assign the values to a vector named stock1 Q2.
k) Store the returns from the second quarter (April, May, June) for stock2.
Assign the values to a vector named stock2 Q2.
l)Find the mean return from stock1 and stock2 for the third quarter (July,
August, and September).
m) When are the returns from stock1 positive?
n) When are the returns from stock2 positive?

``` {r}
stock1 <- c(2000,4000,-500,-1000,7000,7500,-800,5500,-1000) ### Part A

stock2 <- c(3000,-400,-500,1000,-700,7500,1250,6500,1000) ### Part B

stock1 ### Part A
stock2 ### Part B

months <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep" ) ### Part C
months ### Part C

names(stock1) <- months ### Part D
names(stock2) <- months ### Part D

total_monthly <- stock1 + stock2 ### Part E

total_monthly ### Part E

total_stock1 <- sum(stock1) ### Part F
total_stock2 <- sum(stock2) ### Part F

total_9_months <- total_stock1 + total_stock2 ### Part F
total_9_months ### Part F

test_total <- sum(total_monthly) ### Part F
test_total ### Part F

total_stock1 > total_stock2 ### Part G, Total Returns from stock1 DO exceed stock2

stock1_March <- stock1["Mar"] ### Part H
stock1_March ### Part H

stock2_Sept. <- stock2["Sep"] ### Part I
stock2_Sept. ### Part I

Stock1_Q2 <- stock1[c("Apr","May","Jun")] ### Part J
Stock1_Q2 ### Part J

Stock2_Q2 <- stock2[c("Apr","May","Jun")] ### Part K
Stock2_Q2 ### Part K

Stock1_Q3 <- stock1[c("Jul","Aug","Sep")] ### Part L
Stock2_Q3 <- stock2[c("Jul","Aug","Sep")] ### Part L
mean(sum(Stock1_Q3)) ### Part L
mean(sum(Stock2_Q3)) ### Part L

Positive_Stock1 <- which(stock1 > 0) ### Part M
Positive_Stock2 <- which(stock2 > 0) ### Part N
Positive_Stock1 ### Part M
Positive_Stock2 ### Part N

```
### Question 2
a) Create the following matrix using R. Name the matrix product sales.
Product1 sales in AMER = 4000
Product1 sales in APAC = 5000
Product1 sales in EMEA = 4000

Product2 sales in AMER = 6000
Product2 sales in APAC = 5000
Product2 sales in EMEA = 7000

Product3 sales in AMER = 5400
Product3 sales in APAC = 6500
Product3 sales in EMEA = 8500

Product4 sales in AMER = 4000
Product4 sales in APAC = 5500
Product4 sales in EMEA = 4300
b) Name the rows of the matrix as P1, P2, P3, and P4. Name the columns as
AMER, APAC, and EMEA.
c) Calculate the worldwide product sales for each product. Hint: Use rowSums().
Add this information to the product sales matrix as the last column.
Name the column Total Sales. Hint: Use cbind().
d) Calculate the product sales for each product by geos. Hint: Use colSums().
Add this information to the product sales matrix as the last row. Name
the row Total Sales. Hint: Use rbind().
e) Select the APAC sales for all four products.
f) Select the worldwide sales for P3 (across all 3 geos).
g) What are the sales for P3 in APAC.

```{r}
Product1 <- c(4000,5000,4000) ### Part A
Product2 <- c(6000,5000,7000) ### Part A
Product3 <- c(5400,6500,8500) ### Part A
Product4 <- c(4000,5500,4300) ### Part A

products <- c(Product1,Product2,Product3,Product4) ### Part A
product_sales_matrix <- matrix(products, byrow =TRUE, nrow = 4) ### Part A
product_sales_matrix ### Part A

region <- c("AMER","APAC","EMEA") ### Part B
names <- c("P1","P2","P3","P4") ### Part B

colnames(product_sales_matrix) <- region ### Part B
rownames(product_sales_matrix) <- names ### Part B
product_sales_matrix ### Part B

Total_Sales <- rowSums(product_sales_matrix) ### Part C 
Total_Sales ### Part C
product_sales_matrix <- cbind(product_sales_matrix,Total_Sales) ### Part C
product_sales_matrix ### Part C

Total_Sales_Geo <- colSums(product_sales_matrix) ### Part D
product_sales_matrix <- rbind(product_sales_matrix, Total_Sales_Geo) ### Part D
product_sales_matrix ### Part D

product_sales_matrix[, "APAC"] ### Part E

product_sales_matrix["P3", "Total_Sales"] ### Part F

product_sales_matrix["P3","APAC"] ### Part G
``` 
### Question 3
Generate the following table as a data frame in R.
Patient Pulse Rate Body temperature Weight
A 70 98 175
B 68 99 182
C 75 100.8 192
D 59 100.2 165
E 60 99 156
F 72 101.0 155
G 72 100.7 167
H 68 97 169
I 75 101.5 171
Use the dataframe created to answer the following questions. Each question is
independent of the prior questions.
a) Display the first 3 rows of the data frame.
b) Display the last 3 rows of the data frame.
c) Display the structure of the data frame.
d) Sort the data frame by Body temperature in a descending order.
e) Sort the original data frame by Weight in an ascending order.
f) Return all patients whose Body temperature is greater than 100.

g) Return all patients whose Pulse Rate is less than 70.

h) Return all patients whose Weight is more than 170 and Pulse Rate exceeds
70.

i) Return the data frame displaying the last three columns and last three rows.

j) Display only the Pulse Rate column but all the rows.

k) Display only the Pulse Rate and the Weight columns but all the rows.

```{r}

df_Patient <- data.frame("Patient" = c("A","B","C","D","E","F","G","H","I"),
                         Pulse_Rate = c(70,68,75,59,60,72,72,68,75),
                         Body_Temp = c(98,99,100.8,100.2,99,101.0,100.7,97,101.5),
                         Weight = c(175,182,192,165,156,155,167,169,171))
df_Patient

head(df_Patient, n = 3) ### Part A
tail(df_Patient, n = 3) ### Part B
str(df_Patient) ### Part C

temp_descending <- order(df_Patient$Body_Temp, decreasing = TRUE) ### Part D
df_Patient[temp_descending, ] ### Part D

weight_ascending <- order(df_Patient$Weight, decreasing = FALSE) ### Part E
df_Patient[weight_ascending, ] ### Part E

df_Patient[df_Patient$Body_Temp > 100, ] ### Part F

df_Patient[df_Patient$Pulse_Rate < 70, ] ### Part G

df_Patient[df_Patient$Weight > 170 & df_Patient$Pulse_Rate > 70,] ### Part H

df_Patient[7:9, 2:4] ### Part I

df_Patient[1:9, "Pulse_Rate", drop = FALSE] ### Part J

df_Patient[1:9, c(2, 4)] ### Part K




```
