---
title: "Homework 7"
author: "Alberto Pineda"
date: "2023-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
Many airlines use surveys to collect data on customer satisfaction related to
flight experiences. Completing a flight, customers receive an email asking them
to rate a variety of factors, including the reservation process, the check-in
process, luggage policy, cleanliness of gate area, service by flight attendants,
food/beverage selection, on-time arrival, and so on. Suppose that a five-point
scale, with Excellent (E), Very Good (V), Good (G), Fair (F), and Poor (P),
is used to record customer ratings. Assume that passengers on a Delta Airlines
flight from Myrtle Beach, South Carolina, to Atlanta, Georgia, provided the
following ratings for the question, “Please rate the airline based on your over-
all experience with this flight.” The sample ratings are shown on the CSV file
Homework 7 - Question 1. Create a bar chart using either the Base R or the
ggplot2 package and create a visual summary of the overall satisfaction ratings.
What insights can you gain from the bar plot?
```{r}
question1 <- read.csv("homework_7_question_1.csv")

frequency <- table(question1)
frequency <- sort(frequency, decreasing = TRUE)
barplot(frequency, main = "Bar Chart for Customer Experience",
        ylab = "Number of Respondents", col = "blue",
        ylim = c(0, 30))
abline(h = 0, lwd = 2)

## From this Bar Chart, we can see that an overhwelming amount of customers had a Very good experience or excellent experience. As an airline it shows that we have room to improve because most customers did not have an excellent experience. What is good, is that there were few customers in the poor or fair categories, meaning very few customers had negative experiences on this flight. 
```
### Question 2
In order to diversify risk, investors are often encouraged to invest in assets
whose returns have either a negative relationship or no relationship. The data
set Homework 7 - Question 2 shows a portion of the annual returns (in %) for
two assets over the past 20 years. Construct a scatterplot of Return B against
Return A. In order to diversify risk, would the investor be wise to include both of
these assets in her portfolio? The data set is available on the CSV file Homework
7 - Question 2. Create the chart using either the Base R or the ggplot2 package.
```{r}
returns <- read.csv("homework_7_question_2.csv")
View(returns)
return_A <- returns$Return.A
return_B <- returns$Return.B
Year <- returns$Year

plot(x = Year, y = return_A,
     col = "blue",
     xlab = "Year",
     ylab = "Return A",
     main = "Returns over time")
points(Year, return_B,
       col = "green")
legend("topleft", legend = c("Return A", "Return B"),
       col = c("blue", "green"),
       lty = 1,
       cex = 0.5)
## This scatterplot shows that Return B overtime has stayed much mose consistent in its returns while Return A has varied significantly in the amount of returns over time. If an investor wants a consistent and predictable returns it would be wise to only keep Return B. If they want potential for high returns it would be better to keep Return A. 
```
### Question 3
The annual incomes for a sample of 200 first-year accountants were recorded.
Draw a histogram and describe its shape. The data set is available on the CSV
file Homework 7 - Question 3. Create the chart using either the Base R or the
ggplot2 package

```{r}
library(ggplot2)
incomes <- read.csv("homework_7_question_3.csv")
incomes$Incomes <- gsub(",","", incomes$Incomes)
incomes$Incomes <- as.numeric(incomes$Incomes)


ggplot(incomes, aes(x = Incomes)) +
  geom_histogram(aes(color=Incomes, fill=Incomes), 
                 position = "identity", bins = 5, alpha = 0.5) +
  labs(title = 'Frequency of Income',
       x = 'Income',
       y = 'Frequency') +
  scale_color_manual(values = c('#0000FF75', '#B2222275')) +
  scale_fill_manual(values = c('#0000FF75', '#B2222275')) +
  theme(legend.position = "bottom", 
        plot.title = element_text(hjust = 0.5))

### From this histogram we can see that a large majority of individuals fall within the 50,000 income to to 55,000 income. There are very few individuals that report a salary above 70,000. This distribution has a right skew to it and is not a normal distribution
```
### Question 4
The data set shows the median household income (Income in $) and median
house value (House Value in $) for the 50 states. The data set is available on
the CSV file Homework 7 - Question 4.
a) Construct a boxplot for household income and use it to identify outliers, if
any, and comment on skewness.
b) Construct a boxplot for median house value and use it to identify outliers,
if any, and comment on skewness.
Create the charts using either the Base R or the ggplot2 package.

```{r}
home <- read.csv("homework_7_question_4.csv")
median_income <- home$Income
median_value <- home$Value

boxplot(median_income, horizontal = TRUE, 
        main = "Boxplot for Median Household Income",
        xlab = "Income ($)")

boxplot(median_value, horizontal = TRUE, 
        main = "Boxplot for Median Household Value",
        xlab = "Value ($)")

home$z_scores_median_value <- (median_value - mean(median_value))/sd(median_value)
z_scores_median_value <- home$z_scores_median_value
home$Outliers_z <- ifelse(abs(z_scores_median_value) > 3, "Yes", "No")
home

Q3_Median_Value <- quantile(median_value, 0.75)
Q1_Median_Value <- quantile(median_value,0.25)
IQR_Median_Value <- Q3_Median_Value - Q1_Median_Value
home$Outliers_IQR_Median_Value <- ifelse(median_value >= Q3_Median_Value +1.5*IQR_Median_Value, "Yes","No")
home
## In regards to Median Household income, there do not appear to be any outliers according to our boxplot. In regards to Median House value, there are at least 2 values that are outliers. Through additional calculation we can confirm the two outliers in median house value as being California and Hawaii.

```
### Question 5
The monthly value of U.S. exports to China (in $ millions) and imports from
China from 1985 to 2016 was recorded. The data set is available on the CSV
file Homework 7 - Question 5.
a) Draw a line chart of U.S. exports to China.
b) Draw a line chart of U.S. imports from China.
c) Calculate the trade balance and draw a line chart.
d) What do all the charts reveal?
Create the charts using either the Base R or the ggplot2 package.
```{r}
import_exports <- read.csv("homework_7_question_5.csv")
str(import_exports)
exports <- import_exports$U.S..Exports.to.China
imports <- import_exports$U.S..Imports.from.China
Year <- import_exports$Year
View(import_exports)
import_exports$trade_balance <- imports - exports
View(import_exports)
Balance <- import_exports$trade_balance
plot(imports~Year, type="l" , col = "blue", xlab = "Year", 
     ylab = "Billions of Dollars")
lines(exports~Year, type="l", col = "red")
lines(Balance~Year, type="l", col = "green")
title(main = "US & China Trade")
legend("topleft", legend = c("Imports", "Exports","Trade Balance"),
       col = c("blue", "red","green"),
       lty = 1,
       cex = 0.7)
### These charts show us that the amount imported into the US from china is substantially higher than what is exported. From the trade balance we can see it only slightly shifted down due to the sheer difference in imports versus exports. It is clear that while both have increased over time, the trade balance has continued to shift towards imports over time. 
```
### Question 6
Does it matter which shelf a popular brand is placed on? It certainly might,
because we tend to purchase items that are easiest to see. To test this, suppose
that Midway is a large chain of supermarket stores with many stores in many
locations. Midway selects 125 of these stores for an experiment. Specifically, it
selects these particular 125 stores to be as alike as possible, so that store size,
amount of customer traffic, types of customers, and other characteristics are as
similar across stores as possible. Each store stocks cereal in a similar location in
the store on five-shelf displays. In the experiment, 25 randomly selected stores
place a particular popular brand of cereal—we’ll call it Brand X—on the lowest
shelf for a month. Another randomly selected 25 stores place Brand X on the
next-to-lowest shelf, another 25 place it on the middle shelf, another 25 place
it on the next-to-highest shelf, and the final 25 place it on the highest shelf.
Then the number of boxes of Brand X sold is recorded at each of the stores
for the last two weeks of the experiment. (The first two weeks allow customers
to get used to the shelving arrangement.) Does shelf height appear to make a
difference in sales? Create a side-by-side box plot and share your insights. The
data set is available on the CSV file Homework 7 - Question 6. Hint: You will
have to first stack the data. Create the boxplots using either the Base R or the
ggplot2 package.

```{r}
library(tidyverse)
midway <- read.csv("homework_7_question_6.csv")
View(midway)

stacked_midway <- gather(midway, key = "Shelf Location", value = "Value")
View(stacked_midway)

shelf <- stacked_midway$`Shelf Location`
value <- stacked_midway$Value

boxplot(value~shelf,
        main="Sales by Shelf Location",
        horizontal = FALSE,
        xlab="Shelf Location",
        ylab="Sales",
        col="green",
        border="black")
## The boxplots show that there may be a difference between Height. When looking at the mean for "Next to highest" you can see that its mean is higher than "Lowest" by a significant margin. Suggesting that height may play an impact on the sales of items. 

```





































































































































