---
title: "Homework 4"
author: "Alberto Pineda"
date: "2023-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
Create a dataset containing two columns: Method A and Method B. Populate
the column Method A with values 1 through 9 and column Method B with values
10 through 18

```{r}
Method_A <- c(1,2,3,4,5,6,7,8,9)
Method_B <- c(10,11,12,13,14,15,16,17,18)
dataset <- data.frame(Method_A = Method_A, Method_B = Method_B)
dataset
```
### Question 1 A)
Using apply(), find the row means.
```{r}
Row_mean <- apply(dataset, 1,  mean)
Row_mean
```
### Question 1 B)
Using apply(), find the column means
```{r}
Col_mean <- apply(dataset, 2,  mean)
Col_mean
```
### Question 1 C)
Using apply(), find the column sums
```{r}
Col_sum <- apply(dataset, 2,  sum)
Col_sum
```
### Question 1 D)
Using apply(), sort the columns.
```{r}
Col_sort <- apply(dataset, 2,  sort)
Col_sort
```
### Question 1 E)
Using apply(), find the product in each row.
```{r}
Row_Prod <- apply(dataset, 1,  prod)
Row_Prod
```
### Question 1 F)
Print the columns of the dataset using the apply() function
```{r}
apply(dataset, 2, function(x) print(x))
```
### Question 1 G)
Find the length of the dataset columns using the apply() function
```{r}
col_lengths <- apply(dataset,2, length)
col_lengths
```
### Question 1 H)
Use apply() to find the range of numbers in each column of the dataset.
```{r}
col_range <- apply(dataset,2, range)
col_range
```
### Question 2 A)
Use tapply() function to find the mean mileage per number of cylinders.
```{r}
View(mtcars)
tapply(mtcars$mpg, mtcars$cyl, mean)
```
### Question 2 B)
Use tapply() function to find the mean weight per number of cylinders
```{r}
tapply(mtcars$wt, mtcars$cyl, mean)
```
### Question 2 C)
Use tapply() function to find the mean mileage per type of transmission
(auto or manual).
```{r}
tapply(mtcars$mpg, mtcars$am, mean)
```
### Question 2 D)
Use tapply() function to find the mean mileage per number of gears.
```{r}
tapply(mtcars$mpg, mtcars$gear, mean)
```
### Question 3 A)
Split the data set into separate parts split by Education.
```{r}
library(tidyr)
mortgage <- read.csv("homework_4_mortgagelending.csv")
mortgage
splitmortgage <- split(mortgage, mortgage$Education)
splitmortgage
```

### Question 3 B)
Using the lapply function, examine the first 3 rows of every element in the
list.
```{r}
lapply(splitmortgage, head, n=3)
```

### Question 3 C)
Using the lapply function, determine the mean Age, Mortgage, and Income
by writing a custom function
```{r}
myMean <- function(df) {
  mean <- apply(df[, c("Age", "Mortgage", "Income")], 2, mean, na.rm = TRUE)
  rounded_mean <- round(mean, 2)
}

lapply(splitmortgage, myMean)

```
### Question 3 D)
Format the output of part c) above by using the sapply function
```{r}
sapply(splitmortgage, myMean)
```
### Question 4 A)
Create a tibble object named air from the airquality data set.
```{r}
library(tidyverse)
air <- as.tibble(airquality)
str(air)
air
```
### Question 4 B)
Sort the data by the Wind column
```{r}
air <- arrange(air, Wind)
air
```
### Question 4 B)
Remove any rows for which the Ozone column has a missing value.
```{r}
air <- filter(air, Ozone != is.na(air$Ozone))
air
```
### Question 4 C)
Remove the Solar.R column and create a new column containing the ratio
of Ozone to Wind
```{r}
air <- select(air, Ozone, Wind, Temp, Month, Day)
air
air <- mutate(air, OzoneWind_Ratio = Ozone/Wind)
air
```
Question 4 E)
Create a subset of the original airquality data containing just three columns:
Month, Day, and Solar.R. The data should only contain data for June and
July. Name the output solar
```{r}
solar <- subset(airquality, select = c(Month, Day, Solar.R))
solar
solar <- filter(solar, Month == 6:7)
solar
```
### Question 4 F
Merge the air and solar data sets together, retaining all records from the
air data set (that is, a left join)
```{r}
Newair <- left_join(air, solar, copy = F)
Newair
```
Question 4 G)
Calculate the median Ozone value by Month for the merged data
```{r}
tapply(Newair$Ozone, Newair$Month, median)
```
### Question 5 A)
Import the data into R. Then view the dataset using the native viewer in
RStudio
```{r}
Who_hw <- read.csv("homework_4_who.csv")
str(Who_hw)
```
### Question 5 B)
Gather the codes in columns 5 through 61 into a single column named
Code. Store the values in the column Value. Hint: Use the gather() or
the pivot longer() function
```{r}
library(tidyverse)
Who_hw <- pivot_longer(Who_hw, 5:60, names_to = "Code")
Who_hw
```
### Question 5 C)
Split the Code variable into the following variables: new, type, sexage.
Hint: Use the separate() function
```{r}
library(tidyverse)
Who_hw <- separate(Who_hw, Code, into = c("New","Type","sexage"), sep = "_")
Who_hw
```
### Question 5 D)
Split the variable sexage into two – one containing the sex and the second
containing the age. Hint: Use the separate() function.
```{r}
whosep <- separate(Who_hw, sexage, into = c("Sex","Age"), sep = 1)
whosep

```
### Question 5 E)
Spread the type into its own columns. Hint: Use the spread() or pivot wider()
function.
```{r}
library(tidyverse)
whopiv <- pivot_wider(whosep, names_from = "Type")
whopiv
```
### Question 5 F)
Combine the Country and the ISO3 variables into one variable Country ISO.
Hint: Use the unite() function.
```{r}
FinalWho <- unite(whopiv, col = "Country_Iso", iso3, country, sep = "-", remove = T)
FinalWho
```
