---
title: "Homework 3"
author: "Alberto Pineda"
date: "2023-04-26"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
a ) Use the flights.csv file, import it into R using the three packages we
covered in the class, and compare the time it takes to read in this file with
read.csv, read csv, and fread.

```{r}
system.time(flights.csv <- read.csv("homework_3_flights.csv"))
str(flights.csv)
View(flights.csv)
```
### Question 1 a)
```{r}
library(readr)
system.time(flights_csv <- read_csv("homework_3_flights.csv"))
str(flights_csv)
View(flights_csv)
### Very Fast
```
### Question 1 a)
```{r}
library(data.table)
system.time(flights_fread <- fread("homework_3_flights.csv"))
str(flights_fread)
View(flights_fread)

```
### Question 1 B) 
Of the three data.table was the fastests and produced the least amount of errors

### Question 1 C) 
Display the structure of the flights data frame
```{r}
library(data.table)
system.time(flights_fread <- fread("homework_3_flights.csv"))
str(flights_fread)

```
### Question 1 D) 
Display only the first three rows
```{r}
library(data.table)
system.time(flights_fread <- fread("homework_3_flights.csv"))
str(flights_fread)
flights_fread[1:3,]

```
### Question 1 E) 
Display the last 10 rows
```{r}
library(data.table)
system.time(flights_fread <- fread("homework_3_flights.csv"))
str(flights_fread)
tail(flights_fread, n = 10)

```
### Question 1 F) 
Use fread to read the first 1000 lines and only the first 6 columns. Hint:
To get you some practice in reading the help files in R, try to look up help
for the fread to see how to select the first 6 columns. To invoke help on
fread, please type ?fread in your console

```{r}
library(data.table)
system.time(flights_fread <- fread("homework_3_flights.csv", nrows = 1000, select = 1:6))
str(flights_fread)
View(flights_fread)
```
### Question 1 G)
Use the subset command to filter only flights whose carrier is United
Airlines (UA)
```{r}
library(data.table)
system.time(flights_fread <- fread("homework_3_flights.csv"))
str(flights_fread)
subset_flights_fread <- subset(flights_fread, carrier == "UA")
View(subset_flights_fread)
```
### Question 2
Read the facebook.tsv file which is available on the 2UCD platform.
```{r}
library(readr)
facebook <- read_tsv("homework_3_facebook.tsv", show_col_types = FALSE)
str(facebook)
View(facebook)
```
### Question 3
Use this URL to download the San Francisco weather data. http://academic.
udayton.edu/kissock/http/Weather/gsod95-current/CASANFRA.txt
```{r}
library(readr)
url <- "http://academic.udayton.edu/kissock/http/Weather/gsod95-current/CASANFRA.txt"
weather <- read_table(url, col_names = c("Month", "Day", "Year", "Temp"))
weather
View(weather)
```
### Question 4
Which spreadsheets are in this PEW Middle Class Data Excel.xlsx file? Read
in the 3.Median HH income, metro spreadsheet.
```{r}
library(readxl)
excel_sheets("homework_3_pew_middle_class_data_excel.xlsx")

MedianHH <- read_excel("homework_3_pew_middle_class_data_excel.xlsx",
                       sheet ="3. Median HH income, metro")
View(MedianHH)


```


















