---
title: "Homework 1"
author: "Alberto Pineda"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
```
### Question 1
Assign a value 50 to a variable named x. Print out the value of x.

``` {r}
x <- 50
x
```
### Question 2
Assign a value 42 to a variable named my var1. Examine the data type of the
variable my var1. Assign a value ‘‘UCD’’ to a variable named my var2. Ex-
amine the data type of the variable my var2. Assign a value FALSE to my var3.
Examine the data type of the variable my var3

```{r}
var1 <- 42
typeof(var1)

var2 <- "UCD"
typeof(var2)

var3 <- FALSE
typeof(var3)
```
### Question 3
Assign a value of 34 to a variable a. Coerce this variable to a character.

```{r}
a <- 34
a
typeof(a)
a <- as.character(a)
typeof(a)
a
```
### Question 4
Assign a value T to a variable b. Coerce this variable to a character.

```{r}
b <- T
typeof(b)
b <- as.character(b)
b
typeof(b)

```
### Question 5
Coerce the character vector c(‘‘1’’, ‘‘0’’, ‘‘1’’, ‘‘0’’, ‘‘0’’) to log-
ical. Hint: First create this vector and store it in a variable (in an object).
Then use coercion.

```{r}
vec <- c("1","0","1","0","0")
typeof(vec)
num <- as.numeric(vec)
typeof(num)
log <- as.logical(num)
typeof(log)
log
```
### Question 6
Create a vector containing the names of the months – Jan, Feb, ..., Dec.
Coerce this vector to nominal and ordinal factors.

```{r}
vec <- c("Jan","Feb","Mar","Apr","May","June","July","Aug","Sept","Oct","Nov","Dec")
typeof(vec)

nom_vec <- factor(vec)
class(nom_vec)
nom_vec
ord_vec <- factor(vec, ordered = TRUE,
                  levels = c("Jan","Feb","Mar","Apr","May","June","July","Aug","Sept","Oct","Nv","Dec"))
ord_vec
class(ord_vec)

```
### Question 7
Create a variable that stores the value of 3^2*4^1/8

```{r}
answer <- 3^2 * 4^(1/8)
answer
```
### Question 8
Using R, verify that
6x + 42
34.2−3.62 = 29.50556
when x = 2.3.

```{r}
x = 2.3
typeof(x)

answer <- (6*x+42)/(3^(4.2-3.62))
answer
```