---
title: "Midterm Exam"
author: "Alberto Pineda"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
Assign a value 150 to a variable named y. Print out the value of y.
```{r}
y <- 150
y
```
### Question 2
Consider the vector c(1, 3, NA, 8, 5, NA, 3). Determine the sum of the
values in the vector.

```{r}
vec <- c(1, 3, NA, 8, 5, NA, 3)
sum(vec, na.rm = TRUE)
```
### Question 3
Generate the following table as a data frame in R. Then use the ifelse statement
to add a new column named Temp 2 to the data frame. This column should
display High for temperatures of 100◦F and over and display Normal for other
values.
Patient Pulse Rate Body temperature Weight
A 70 98 175
B 68 99 182
C 75 100.8 192
D 59 100.2 165
E 60 99 156
F 72 101.0 155
G 72 100.7 167
H 68 97 169
I 75 101.5 171

```{r}
df_Patient <- data.frame("Patient" = c("A","B","C","D","E","F","G","H","I"),
                         Pulse_Rate = c(70,68,75,59,60,72,72,68,75),
                         Body_Temp = c(98,99,100.8,100.2,99,101.0,100.7,97,101.5),
                         Weight = c(175,182,192,165,156,155,167,169,171))
df_Patient$Temp_2 <- ifelse(df_Patient$Body_Temp > 100, "High", "Normal")
df_Patient
```
### Question 4
Use this URL to download the Sacramento weather data. https://academic.
udayton.edu/kissock/http/Weather/gsod95-current/CASACRAM.txt.
```{r}
library(readr)

url <- "https://academic.udayton.edu/kissock/http/Weather/gsod95-current/CASACRAM.txt"
weather <- read_table(url, col_names = c("Month", "Day", "Year", "Temp"))
View(weather)
```
### Question 5
The Wall Street Journal CEO Compensation Study analyzed CEO pay for many
U.S. companies with fiscal year 2008 revenue of at least $5 billion that filed their
proxy statements between October 2008 and March 2009. The data are in the
CSV file Question 5.

Part a)Read the CSV file into the R environment.

```{r}
Comp <- read.csv("2114e0b0-829f-488c-8706-ba2b1a1f796d.csv")
View(Comp)
```
### Part B)
Create a new variable that is the sum of Salary 2008 and Bonus 2008.
Name this new variable Total Compensation. Add this new variable to
the data frame created in part a) above.
```{r}
Comp$Salary_Clean <- as.numeric(gsub("\\$|,","", Comp$Salary.2008))
Comp$Bonus_Clean <- as.numeric(gsub("\\$|,","", Comp$Bonus.2008))
View(Comp)
Comp$TotalCompensation <- rowSums(Comp[,c("Bonus_Clean","Salary_Clean")])
Comp$TotalCompensationFormatted <- paste("$", format(Comp$TotalCompensation, big.mark = ",", decimal.mark = "."), sep = "")
View(Comp)
```
### Part C)
Display the first 3 rows of the data frame
```{r}
head(Comp, n = 3)
```
### Part D)
```{r}
tail(Comp, n = 3)
```
### Part E)
Display the structure of the data frame
```{r}
str(Comp)
```
### Part F)
How many rows are in the data frame?
```{r}
nrow(Comp)
```
### Part G)
Sort the data frame by Total Compensation in descending order. Who
earned the largest compensation?

```{r}
TotalCompDescending <- order(Comp$TotalCompensation, decreasing = T)
Comp[TotalCompDescending,]
### Laurence D. Fink earned the largest compensation at $17,649,500
```
### Part H)
Retrieve all the rows where the Total Compensation is greater than $10
million
```{r}
Comp[Comp$TotalCompensation > 10000000,]
```
### Part I)
Return the rows for the CEO compensation for the CEOs in the Financials
sector
```{r}
Comp[Comp$Company.Type == "Financials",]
```
### Part J)
Return the rows for the CEO compensation for the CEOs in the Health
Care or the Technology sectors.
```{r}
Comp[Comp$Company.Type == "Technology" | Comp$Company.Type == "Health Care",]
```
### Part K)
Sort the data frame by Company Type ascending and Total Compensation
descending.

```{r}
Comp[order(Comp$Company.Type, -Comp$TotalCompensation),]
View(Comp[order(Comp$Company.Type, -Comp$TotalCompensation),])
```




